---
title: "Visualizations"
author: "She Chee Yee (A0240383L)"
date: "2023-04-29"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(lubridate)
library(tidyverse)
library(stringr)

```



```{r }
volcano = read.csv("../data/volcano_practice.csv")
region = read.csv("../data/continents_practice.csv")

```

## Including Plots
How many countries are there in this data set?
How many unique types of volcanoes are there in the data?
What is the second most common volcano type in the data?

```{r }
volcano %>% distinct(country) %>% count()

volcano %>% mutate(type = str_remove(primary_volcano_type, "\\(s\\)|\\(es\\)|\\?")) %>%
                     distinct(type) %>% count()

volcano %>% mutate(type = str_remove(primary_volcano_type, "\\(s\\)|\\(es\\)|\\?")) %>% 
  group_by(type) %>% count(sort = TRUE)

```

```{r}
volcano = volcano %>% mutate(type = str_remove(primary_volcano_type,"\\(s\\)|\\(es\\)|\\?"))

volcano %>% mutate(type = reorder(type, elevation, median)) %>%
  ggplot(aes(x = type, y = elevation)) +
  geom_boxplot() +
  coord_flip()



```
Next, let us examine the number of volcanoes across continents.
The file continents_practice.csv classifies countries into
continents. We’ve read it in as region.
Fill in the blanks in the code to create the bar chart below.

2 ways to do this.
1. use geom_bar(), aes is Continent, geom_text with `label = ..count.., stat = "count"

2. summarize the data with group_by, geom_col(x = continents, y = count, label = count) then geom_text()

```{r}
continents= read.csv("../data/continents_practice.csv")

volcano =volcano %>% left_join(continents, by = c("country" = "Entity")) 



ggplot(data = volcano, aes(Continent)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count")

v1 =volcano %>% group_by(Continent) %>%
  summarize(count = n())

ggplot(data = v1, aes(Continent, count, label = count)) +
  geom_col() +
  geom_text(nudge_y = 10.0)

v1 %>% mutate(pct = count/sum(count) * 100)



```





The company wants to figure out when people begin Christmas
shopping, so that they can start preparing for it.
▶ Write the R code to create an object named qn1 that contain the
number of Christmas-related items sold in each month.
Specifically, look for the following keywords in the Description
column: CHRISTMASS, XMAS, REINDEER, or SANTA

```{r}


retail = readRDS("../data/retail_practice.rds")%>%
  mutate(InvoiceDate = as.Date(InvoiceDate),
         month = month(InvoiceDate, label = TRUE, abbr =TRUE),
         year = year(InvoiceDate), 
         day = day(InvoiceDate))


qn1 = retail %>% filter(str_detect(Description, "CHRISTMAS|XMAS|REINDEER|SANTA")) %>%
  group_by(year, month) %>% summarize(n = n(), .groups = "drop")

qn1.1 = retail %>% filter(str_detect(Description, "CHRISTMAS|XMAS|REINDEER|SANTA")) %>%
  count(year, month)

qn1 %>% head(4)
qn1.1 %>% head(4)



```



```{r}

retail %>% group_by(year, month) %>% distinct(day) %>% count()

qn2 = retail %>% 
  mutate(InvoiceDate = as_date(InvoiceDate),
         month = substr(InvoiceDate, 1, 7)) %>%
  # Compute number of unique days in each month and monthly revenue
  group_by(month) %>%
  summarize(days = n_distinct(InvoiceDate),
            rev = sum(UnitPrice*Quantity)/1000) %>%
  ungroup()

qn2 %>% filter(month != "2011-12") %>% summarize(mean_days = mean(days))




ggplot(data = qn2, aes(x = factor(month), y = rev)) +
  geom_point(size = 6) +
  geom_line(aes(group = 1)) +
  geom_text(aes(label = days), color = "white", size = 2.7) +
  annotate(geom = "text", x = 12.6, y = 53.5, label = "days") +
  labs(title = "Monthly revenue in thousands of pounds, all items",
       x = "Month", y = "Revenue in thousands of pounds",
       subtitle = "Number of unique transaction days in each month shown in text.") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30))

```
Notice that 2011 Dec only have 8 days, we can try to increase the number of days and quantify by scaling the number of days to average days per month in the dataset.





```{r}
library(nycflights13)
library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)

flights %>% filter(month == 12, origin == "JFK") %>% group_by(dest) %>% count(sort = TRUE)

flights %>% group_by(carrier) %>%
  summarize(count = n(), avg_dist = sum(distance)/count) %>%
  arrange(desc(avg_dist))

flights %>% filter(arr_delay > 0, !is.na(arr_delay)) %>%
  group_by(month, day) %>%
  summarize(avg_delay = sum(arr_delay) / n()) %>%
  arrange(desc(avg_delay))


flights %>% filter(month == 7, day == 4) %>% 
  left_join(planes, by = c("tailnum")) %>% 
  filter(!is.na(dep_time)) %>%
  summarize(avg_seats = round(mean(seats, na.rm = TRUE), 2)) %>% pull(avg_seats)

flights %>% mutate(dep_hour = as.integer(sched_dep_time / 100)) %>%
  group_by(dep_hour) %>%
  mutate(over = case_when(dep_delay > 15 ~ 1,
                          TRUE ~ 0)) %>%
  summarize(p = sum(over) / n()) %>%
  arrange(desc(p))
# What plane traveled the most times in 2013? Plot the number of trips per week over the year for that plane.
plane = flights  %>% left_join(planes, by = c("tailnum" = "tailnum")) %>%
  filter(year.x == 2013, !(is.na(dep_time))) %>% 
  group_by(tailnum) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>% 
  top_n(1, count) %>% pull(tailnum)

most = flights %>% filter(tailnum == plane, !(is.na(dep_time)))
most_trip = flights %>% filter(tailnum == plane, !(is.na(dep_time))) %>%
  unite(col = date,year,  month, day, sep = "-") %>%
  mutate(date = as.Date(date),
         week = week(date)) %>%
  group_by(week) %>%
  summarize(count = n())

ggplot(data = most_trip, aes(week, count)) + 
  geom_point(color = "red3", size = 3) + 
  geom_line() + 
  labs(x = "Week of the year", y = "", title = "Trips per week for N725MQ in 2013") + 
  theme_minimal()
  
colnms <- colnames(most)

# filter
most %>%
  filter_at(vars(all_of(colnms)), any_vars(is.na(.)))
```

scatterplot

```{r}
data(mpg)
str(mpg)

df = mpg %>% mutate(trans = str_sub(trans, 1, 4))

df1 = mpg %>% mutate(trans = case_when(str_detect(trans, "^auto") ~ "auto",
                                       str_detect(trans, "^manual") ~ "manual",
                                       TRUE ~ trans))

ggplot(data = df1, aes(x = displ, y = hwy, label = trans)) +
  geom_point(size = 2,  alpha = 1) 

ggplot(data = df1, aes(x = displ, y = hwy, label = trans)) +
  geom_point(aes(shape = class),position = "jitter")  + 
  scale_shape_manual(values = 1:7)

ggplot(data = cars, aes(speed, dist)) + 
  geom_point(aes(size = 2.0,alpha = 0.4), show.legend = FALSE) +
  labs(title = "Relationship between Speed and Braking", 
       ylab = "Stopping distance (ft)",
       xlab = "Speed (mph)")

```

Histogram

```{r}
heights = read.csv("../data/heights.csv", header= TRUE, stringsAsFactors = TRUE)

ggplot(data = heights) + 
  geom_histogram(aes(x = earn/1000, y = ..density.., fill = sex),  
                 binwidth = 10, color = "white", boundary = 0) 

ggplot(data = heights) + 
  geom_histogram(aes(x = earn/1000, y = ..density.., fill = sex),  
                 binwidth = 10, color = "white", boundary = 0, position = "dodge") +
  scale_fill_discrete(name = "Gender", labels = c("Female", "Male"))

ggplot(data = heights) + 
  geom_density(aes(x= earn/1000, fill = sex), alpha = 0.2)

```
CDF

```{r}

heights %>%
  filter(sex == "female") %>%
  sample_n(20) %>%
  ggplot(aes(x = earn/1000)) + 
  stat_ecdf()

heights %>%
  ggplot(aes(x = earn/1000, col = sex)) +
  stat_ecdf()


```
normal dist

```{r}

ggplot(data = data.frame(x = seq(-5, 5)), aes(x)) +
stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1)) +
labs(x = "", y = "", title = "A standard normal distribution")

heights_f = heights %>% filter(sex == "female")

heights_f %>%
  ggplot(aes(height)) + 
  geom_density(fill = "lightblue", alpha = 0.5) +
  stat_function(fun = dnorm, args = list(mean(heights_f$height), sd(heights_f$height)))


```
line graphs

```{r}


UN_data = read_excel("../data/UNESCAP_population_2010_2015.xlsx", sheet = 3)
head(UN_data)

pop1 = UN_data %>% gather(Y2010:Y2015, key = years, value = population) %>%
  mutate(years = as.integer(str_sub(years, 2, 5))) %>%
  filter(e_fname %in% c("Singapore", "Malaysia", "Cambodia",
  "Thailand", "Viet Nam", "Lao PDR",
  "Brunei Darussalam"))

ggplot(data = pop1, aes(years, population)) +
  geom_line(aes(color = e_fname))

end_pts = pop1 %>% filter(years == 2015) %>%
  mutate(fname = recode(e_fname, "Brunei Darussalam" = "Brunei"))

ggplot(data = pop1, aes(years, population, group = e_fname)) + 
  geom_line() +
  geom_point(data = end_pts, aes(years)) + 
  geom_text(data = end_pts, aes(label = fname), hjust = "left", nudge_x = 0.1) +
  geom_vline(xintercept = 2011, linetype = "dashed") +
  geom_hline(yintercept = 1000, linetype = "dashed") +
  geom_abline(slope = 1)

```
reconstruction of graph

```{r}
library(dslabs)
library(ggthemes)

df = murders
head(df)

df %>% distinct(state)
df %>% count(state)

r = df %>% summarize(rate = sum(total)/ (sum(population) / 1e6)) %>%
  pull(rate)

df %>% ggplot(aes(x = population/1e6, y = total)) +
  geom_point(aes(color = region), size = 3,  show.legend = FALSE) +
  geom_text(aes(label = abb), nudge_x = 0.07) +
  scale_x_log10() + 
  scale_y_log10() +
  labs(title = "US Gun Murders in 2010",
    x = "Population in millions (log scale)",
    y = "Total number of murders (log scale)") +
  geom_abline(slope = 1, intercept = log10(r), lty = 2, color = "darkgrey") + 
  theme_economist()
  
    

```



bar charts

```{r}

budget_cat = c("Manpower", "Asset", "Other") # create data frame
amount = c(519.4, 38.0, 141.4)
op_budget = data.frame(budget_cat, amount)
barplot(op_budget$amount, border = NA, names.arg = op_budget$budget_cat)

ggplot(data = op_budget, aes(budget_cat, amount)) +
  geom_col()

op_budget %>% mutate(budget_cat = reorder(budget_cat, -amount)) %>%
  ggplot(aes(budget_cat, amount)) + 
  geom_col()

```














```{r}

imda_data = readRDS("../data/imda.rds")
str(imda_data)

young_adults = imda_data %>% filter(age == "20-29", year == 2015) %>%
  mutate(pct = as.numeric(ever_used),
         media_activity = reorder(media_activity, pct))

ggplot(data = young_adults) +
  geom_col(aes(media_activity, pct, fill = media_activity)) + 
  coord_flip()


tv = imda_data %>%
  filter(media_activity %in% c("Watch MediaCorp TV",
                               "Watch StarHub TV", "Watch Singtel TV")) %>%
  mutate(pct = as.numeric(ever_used),
         age = reorder(as.factor(age), as.numeric(substr(age, 1, 2))))


p1 = ggplot(tv, aes(x = year, y = pct)) +
  geom_col(aes(fill = media_activity),
           position = "dodge") +
  facet_wrap(~ age) +
  labs(title = "Percentage of Young Adults ...",
       x = "Activity", y = "Percentage", fill = "Activity") +
  theme(legend.position = "bottom")


```





smoothed line & loess

```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 5)) +
  labs(x = "Engine Displacement (l)", y = "Highway Miles per Gallon")

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "loess", span = 0.2) +
  labs(x = "Engine Displacement (l)", y = "Highway Miles per Gallon")


```

loess by group


```{r}

ggplot(mpg, aes(x = displ, y = hwy, col = drv)) +
  geom_point(position = "jitter") +
  geom_smooth(aes(linetype = drv),
              method = "loess", show.legend = FALSE) +
  labs(x = "Engine Displacement (l)", y = "Highway Miles per Gallon") +
  scale_color_discrete(name = "Drive type",
  labels = c("Four-wheel", "Front-wheel", "Rear-wheel")) +
  theme(legend.position = "top")
```

rug plot

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(position = "jitter") +
  geom_rug(position = "jitter") +
  labs(x = "Engine Displacement (l)", y = "Highway Miles per Gallon")


```

boxplot

```{r}

df = mpg %>% mutate(class = reorder(class, hwy, FUN = median))

ggplot(df, aes(x = class, y = hwy)) +
  geom_boxplot(outlier.color = "red",  outlier.shape = 1) +
  labs(x = "Class of Car", y = "Highway Fuel Efficiency")




```


remove outliers

```{r}
heights2 = heights %>%
  group_by(sex) %>%
  mutate(max_earn = quantile(earn, 0.75) + 1.5 * IQR(earn),
  min_earn = quantile(earn, 0.25) - 1.5 * IQR(earn)) %>%
  filter(earn <= max_earn & earn >= min_earn) %>% ungroup()

ggplot(data = heights2) +
geom_boxplot(aes(x = earn, y = sex))


```

2D Histograms
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_bin2d(binwidth = c(0.5, 5)) +
  scale_fill_gradient(name = "Count", low = "white", high = "red") +
  labs(title = "2D Bins",
       x = "Engine Displacement (l)", y = "Highway Miles per Gallon")


ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_hex(binwidth = c(0.5, 5)) +
  scale_fill_gradient(name = "Count", low = "white", high = "red") +
  labs(title = "Hexagonal Bins",
       x = "Engine Displacement (l)", y = "Highway Miles per Gallon")


```


```{r}
library(ggrepel)
data(gapminder)

summary(gapminder)

gapminder %>% filter(!is.na(gdp)) %>%
  group_by(year) %>%
  count()

df = gapminder %>% filter(year == 2011) %>%
  mutate(gdpc = gdp / population) %>%
  filter(!is.na(gdpc))
  

ggplot(data = df, aes(gdpc, life_expectancy)) + 
  geom_point(aes(color = continent, size = population)) + 
  scale_x_log10() +
  geom_text_repel(aes(label = country), show.legend = FALSE, size = 2) +
  guides(size = "none")

```


Top streaming services in the US: There are only a few data points on the graph. We
can manually create the data frame.

```{r}
services = c("Netflix", "PrimeVideo", "hulu", 
             "Disney", "AppleTV", "peacock", "HBO", "Paramount", "others")
level =  c("Netflix", "PrimeVideo", "hulu", 
             "Disney", "AppleTV", "peacock", "HBO", "Paramount", "others")

streaming = tibble::tribble(
  ~service, ~`2020`, ~`2021`,
  "Netflix", 29, 20,
  "Prime", 21, 16,
  "Hulu", 16, 13,
  "Disney", 12, 11,
  "Apple", 4, 5,
  "Peacock", 0, 5,
  "HBO", 3, 12,
  "Paramount", 2, 3,
  "Others", 13, 15
)


y_2020 = c(29,21,16,12,4,0,3,2,13)
y_2021 = c(20,16,13,11,5,5,12,3,15)

df = data.frame(services, y_2020, y_2021)

df = df %>% gather(`y_2020`:`y_2021`, key = "year", value = "pct") %>%
  mutate(year = (str_sub(year, 3, 7))) %>%
  mutate(services = factor(services, ordered = TRUE,
                           levels = level))

ggplot(data = df, aes(services, pct, label = paste0(pct, "%"), fill = year)) +
  geom_col(position = "dodge") + 
  geom_text(aes(services, pct + 1, group = year), position = position_dodge(width = 1)) + 
  scale_fill_manual(values = c("2020" = "red3", "2021" = "black")) + 
  scale_y_continuous(labels = scales::label_number(suffix = "%")) + 
  labs(x = "", y= "", title = "US Streaming Market Share") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "top")

```
improvement

```{r}
growth = df %>%
  group_by(services) %>%
  mutate(pct_change = pct - lag(pct)) %>%
  na.omit() 


ggplot(data= growth) + 
  geom_col(aes(services, pct_change, fill = pct_change > 0)) + 
  coord_flip() + 
  scale_fill_manual(values = c(`TRUE` = "darkcyan", `FALSE` = "red3"))

```



How Chinese New Year compares with Thanksgiving: cny_thanksgiving.xls

```{r}


cny = read_excel("../data/cny_thanksgiving.xlsx") %>%
  mutate(Population = case_when(Country == "China" ~ 1339724852,
                                Country == "USA" ~ 307745538),
         Per_capita = Amount/Population,
         Amount = Amount / 1e9) %>%
  filter(Category != "Population Census 2010") %>%
  rename(total = Amount) %>%
  gather(total, Per_capita, key = "mertics", value = "amount")

ggplot(data = cny %>% filter(mertics == "Per_capita"), aes(Country, amount, fill = Country)) + 
  geom_col(show.legend = FALSE)+ 
  facet_wrap(~ Category, scales = "free_y", labeller = label_wrap_gen()) + 
  scale_fill_manual(values = c("China" = "red3", "USA" = "blue")) + 
  labs(x = "" , y = "", fill = "",
       title = "How does Chinese New Year compare to Thanksgiving?") + 
  theme_minimal()


```







Monthly cereal consumption: personal_consumption_expenditures.xlsx.

```{r}
cereal = read_excel("../data/personal_consumption_expenditures.xlsx") %>%
  filter(`Sub-Category` == "Cereals") %>%
  rename(subcat = `Sub-Category`,
         amount = `Millions of Dollars`) %>%
  mutate(amount = amount /1000)

ggplot(data = cereal, aes(x = Month, y = amount)) +
  geom_line(linewidth = 2) + 
  geom_area(fill = "cornflowerblue", alpha = 0.2) + 
  labs(x= "", y = "", title = "Well that sure was unusual",
       subtitle = "Monthly US personal consumption expenditures on cereals",
       caption = "Source: US Burea") + 
  theme_minimal()



```
graph 2
```{r}
covid = cereal %>% mutate(year = year(Month),
                          month = month(Month, label = TRUE, abbr = TRUE)) %>%
  filter(year == 2020 | year == 2019) %>%
  select(amount:month) %>%
  spread(key = year, value = amount)


ggplot(data = covid, aes(month, amount)) +
    geom_segment(aes(x = month, y = `2020`, xend = month, yend = `2019`), color = "grey") + 
  geom_point(data = covid, aes(x = month, y = `2019`), color = "grey", alpha = 0.5, size = 5) + 
  geom_point(data = covid, aes(x = month, y =`2020`), color = "indianred4", size = 5) + 
  labs(x = "", y= "Expenditure in billions of dollars",
       title = "Consumption of Cereals during the pandemic, 2020 (red) 2019 (grey)", 
       caption = "Source: US Burea of Economic Analysis")


```



Who do we spend time with in our life: time-spent-with-relationships-by-age-us.csv.


```{r}
time = read.csv("../data/time-spent-with-relationships-by-age-us.csv")
time = time %>% rename(alone = 4,
                friends = 5,
                children = 6,
                family = `Time.spent.with.with.parents..siblings.and.other.family..by.age.of.respondent..United.States.`, 
                partner = `Time.spent.with.partner..by.age.of.respondent..United.States.`,
                coworkers = `Time.spent.with.coworkers..by.age.of.respondent..United.States.`,)

time = time %>% gather(`alone`:`coworkers`, key = "type", value = "minutes") %>%
  filter(Year <= 80)

labels = time %>% filter(Year == 80 )

ggplot(data = time, aes(Year, minutes, color = type), show.legend = FALSE) +
  geom_point(show.legend = FALSE) + 
  geom_line(show.legend = FALSE) + 
  geom_text_repel(data = labels, aes(label = type), show.legend = FALSE, nudge_x = 3.5) + 
  scale_y_continuous(labels = scales::label_number(suffix = " mins"),
                     breaks = seq(0,400,100)) + 
  scale_x_continuous(breaks = c(15,20,30,40,50,60,70,80), limits = c(15,90)) + 
  labs(x = "Age", y = "", 
       title = "Who Americans spend their time with, by age") + 
  scale_x_continuous(guide = "none") + # this removes the legend 
  theme_minimal() + 
  labs(x = "", y= "lorem ipsum",
       title = "lorem ipsum", 
       caption = "Source: US Burea of Economic Analysis")
  
  


```


NBA plots

```{r}

library(ggrepel)
nba = read.csv("../data/NBA_Team_stats.csv") %>%
  separate(Year, into = c("start_year", "end_year"), convert = TRUE) %>%
  filter(start_year >= 2010) %>%
  select(`No`:`Pf`, start_year) %>%
  gather(`Pts`:`Pf`, key = "type", value = "values")

pts = nba %>% filter(type == "Pts") 
end_labels = pts %>% filter(start_year == 2021) %>%
  top_n(10, values)

top_10 = end_labels %>% pull(Team)

pts = pts %>% filter(Team %in% top_10)

ggplot(data = pts, aes(start_year, values, color = Team, show.legend = FALSE)) +
  geom_line(aes(groups = Team), show.legend = FALSE) +
  geom_text_repel(data = end_labels, aes(start_year, values, label = Team), show.legend = FALSE, 
                  nudge_x = 2) +
  scale_x_continuous(labels = scales::label_number(suffix = " season")) + 
  scale_y_continuous(breaks = seq(100,130,5)) +
  labs(x = "Points", y = "Seasons", title = "NBA Average points for each team (2010-2021)" ) + 
  theme_minimal()

```
nba players

```{r}

players = read.csv("../data/NBA_Player_Stats.csv") %>%
  separate(Year, into = c("start", "end")) %>%
  mutate(start = as.numeric(start), 
         end = as.numeric(end)) %>%
  filter(start >= 2010)

players %>% distinct(Player)

allstars_2023 = c("Giannis Antetokounmpo", "Lauri Markkanen", "Donovan Mitchell", "Ja Morant",
                  "Jayson Tatum", "LeBron James", "Luka Dončić", "Joel Embiid", "Kyrie Irving",
                  "Nikola Jokić")

global_type = "AST"
allstars = players %>% filter(Player %in% allstars_2023) %>%
  gather(`G`:`PTS`, key = "type", value = "values")

end_labelss = allstars %>% filter(end == 2022, type == global_type) %>% arrange(desc(values))

allstars_df = allstars %>% mutate(Player = factor(Player, levels = end_labelss %>% pull(Player)))

ggplot(data = allstars_df %>% filter(type == global_type), aes(start, values, color = Player)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(labels = scales::label_number(suffix = "season")) +
  labs(x = "Points", y = "Seasons",
       title = "Average AST per season for the 2023 All-Star teams from 2010 season onwards") +
  theme_minimal()
```

jimmmy butler

```{r}

butler = tibble::tribble(
  ~season, ~`reg`, ~`playoffs`,
  "2022-2023", 23.0, 36.5,
  "2021-2022", 21.4, 27.4,
  "2020-2021", 21.5, 14.5,
  "2019-2020", 33.8, 38.4,
  "2018-2019", 33.2, 35.1,
  "2017-2018", 22.2, 15.8,
  "2016-2017", 37.0, 39.8,
  "2014-2015", 38.7, 42.2,
  "2013-2014", 38.7, 43.5,
  "2012-2013", 26.0, 40.8,
)


colors = c("reg" = "grey", "playoffs" = "indianred4")

ggplot(data = butler, aes(season, points)) + 
  geom_segment(aes(x = season, y = `reg`, xend = season, yend = `playoffs`), color = "grey")+
  geom_point(data = butler, aes(x = season, y = `reg`, color = "reg"), size = 5) +
  geom_point(data = butler, aes(x = season, y = `playoffs`, color = "playoffs"), size = 5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "Seasons", y= "Average points",
       title = "Scoring Performance of Jimmy Butler",
       caption = "Source: hoopsstats.com",
       color = "Legend") +
  scale_color_manual(values = colors,  labels = c("Playoffs", "Regular Season")) + 
  theme_clean() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 9, face = "italic"),
        plot.caption = element_text(face = "italic"),
        legend.position = "top",
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9))
butler = butler %>% gather(`reg`:`playoffs`, key = "period", value = "points")

```

BASE R

```{r}

a1 = seq(13, 21)
a1

a2 = seq(13,21, by = 2)
a2

a1 + a2

sum(1:100) == 100* (100+1)/2

data(airquality)
summary(airquality)
class(airquality)
str(airquality)

sum(is.na(airquality ))

which(is.na(airquality))

df = na.omit(airquality)

df_summer = df[df$Month %in% c(5,6),]
df_fall = df[df$Month %in% c(7, 8, 9),]

df$Season = ifelse(df$Month %in% c(5,6), "Summer", "Fall")
df$Season = factor(df$Season)

summary(df$Temp)

str(df)

tapply(df$Temp, df$Season, summary)

df$Month[df$Month == 5] = "May"
df$Month[df$Month == 6] = "June"
df$Month[df$Month == 7] = "July"
df$Month[df$Month == 8] = "August"
df$Month[df$Month == 9] = "September"
df$Month = factor(df$Month, levels = c("May", "June", "July", "August","September"))


boxplot(df$Temp ~ df$Month, xlab = "Month", ylab = "Temperature",
          main = "Monthly temperature in New York")

ggplot(data = df, aes(Month, Temp)) + 
  geom_boxplot() +
  labs(x = "Month", y = "Temperature", title = "Monthly temperature in New York")

```
LUBRIDATE

```{r}

shanghai = read.csv("../data/Shanghai.csv")

head(shanghai, 3)


shanghai$Date = paste0(shanghai$Year, "-", shanghai$Month)

shanghai$Date = ym(shanghai$Date)
shanghai

shanghai_2012 = shanghai[shanghai$Year == 2012,]

plot(shanghai_2012$Month, shanghai_2012$Temp, xlab = "Month", 
     ylab = "Temperature", type = "b", cex = 01, pch = 19,
     main = "Monthly temperature in Shanghai 2012") 

```
FUNCTIONS


```{r}
# Q1.1 pass_check
pass_check = function(grade) {
  
  # Check if the input is numeric
  # If not, stop the current execution and print an error
  if(!is.numeric(grade)) {
    stop("The input is not numeric.")
  } 
  
  if (grade >= 5){
    print("pass")
  } else {
    print("fail")
  }
  
}

factor_convert = function(df) {
  
  for(i in 1:ncol(df)) {
    
    var = df[, i]
    
    if(is.character(var)) {
      df[, i] = as.factor(var)
    }
  }
  
  return(df)
  
}

select(df, where(is.factor)) 

df[,sapply(df, is.factor)]


```



RDS


```{r}

hawk = readRDS("../data/hawker_ctr_raw.rds")
hawk = hawk[[1]][-1]

all_names = sapply(hawk, function(x) x$NAME)

compute_dist = function(name1, name2) {
  
  id_1 = which(all_names == name1)
  id_2 = which(all_names == name2)
  
  coord_1 = as.numeric(str_split(hawk[[id_1]]$XY, ",")[[1]])
  coord_2 = as.numeric(str_split(hawk[[id_2]]$XY, ",")[[1]])
  
  dist = sqrt(sum((coord_1 - coord_2)^2))
  dist
}

dist_example = compute_dist("Tanglin Halt Market", "Taman Jurong Market & Food Centre")

hawk_combn = combn(all_names, 2)
compute_dist_v = Vectorize(compute_dist)
all_dist = compute_dist_v(hawk_combn[1, ], hawk_combn[2, ])
dist_df = data.frame(hawker1 = hawk_combn[1, ],
                     hawker2 = hawk_combn[2, ],
                     dist = all_dist, stringsAsFactors = FALSE)

dist_df


```


TIDY DATA
```{r}

unicef_hiv_tidy = read_excel("../data/unicef_hiv_data.xlsx", 
                             range = "A1:O219", na = c("-", "<0.01")) %>%
  filter(!(row_number() %in% c(1,2))) %>%
  gather(`2021`:`2010`, key = "years", value = "values") %>%
  mutate(values = as.numeric(values)) %>%
  spread(key = indicator, value = values) %>%
  rename(incidence = 4,
         deaths = 5,
         country = 1)



global_evolution = unicef_hiv_tidy %>%
  group_by(years) %>%
  summarize(total = sum(incidence, na.rm = TRUE)) %>%
  ungroup()

ggplot(data = global_evolution, aes(years, total)) + 
  geom_line(aes(group = 1))


iso_unicef = read.csv("../data/iso_unicef_region.csv")

overall = unicef_hiv_tidy %>% left_join(iso_unicef, by = c("country" = "Country.Region")) %>% 
  filter(!(is.na(UNICEF.Region)))

global_average = overall %>%
  group_by(years) %>%
  summarize(avg = mean(incidence, na.rm = TRUE)) %>%
  ungroup()


africa = overall %>% filter(UNICEF.Region == "Eastern and Southern Africa") %>%
  group_by(years) %>%
  summarize(avg = mean(incidence)) %>% 
  ungroup()

ggplot(data = africa) + 
  geom_line(aes(years, avg, group = 1), color = "indianred4", linewidth = 1.5) + 
  geom_line(data = global_average, aes(years, avg, group = 1), color = "grey", linewidth = 1.5) +
  annotate(geom = "text", x = 10, y = 1.2, label = "Africa") +
  annotate(geom = "text", x = 10, y = 0.63, label = "World") + 
  labs(y = "Average incidence rate",
  title = "HIV incidence rates in Eastern and Southern Africa to the global average")

```



JOINS

```{r}

themes = read.csv("../data/archive/themes.csv")
sets = read.csv("../data/archive/sets.csv")
inventory_sets = read.csv("../data/archive/inventory_sets.csv")
colors = read.csv("../data/archive/colors.csv")
inventories = read.csv("../data/archive/inventories.csv")
inventory_parts = read.csv("../data/archive/inventory_parts.csv")
parts = read.csv("../data/archive/parts.csv")
part_categories = read.csv("../data/archive/part_categories.csv")

part_cat = part_categories %>% 
  filter(str_detect(name, "Bricks"))

themes %>% anti_join(sets, by = c("id" = "theme_id")) %>% count()

brick_parts = inventory_parts %>% 
  left_join(colors, by = c("color_id" = "id")) %>%
  left_join(parts, by = c("part_num" = "part_num")) %>%
  left_join(part_cat, by = c("part_cat_id" = "id")) %>%
  rename(part_name = 2,
         part_cat_name = 3,
         color_name = 4) %>%
  na.omit()



inv_parts_col = inventory_parts %>% left_join(colors, by = c("color_id" = "id"))
parts_cat = parts %>% left_join(part_categories, by = c("part_cat_id" = "id"), 
                                suffix = c("_parts", "_cats")) %>%
  filter(str_detect(name_cats, "Bricks"))

table = inv_parts_col %>% left_join(parts_cat, by = c("part_num" = "part_num")) %>%
  filter(!(is.na(name_cats))) %>%
  select(part_num, name_parts, name_cats, name, rgb)


pop_themes = themes %>% 
  left_join(sets, by = c("id" = "theme_id"), suffix = c("_themes", "_sets")) %>% 
  filter(str_detect(name_themes, "Star Wars")) %>%
  group_by(year) %>%
  summarize(numSets = n_distinct(set_num),
            avg = ceiling(mean(num_parts)))

nrow(sets)
nrow(inventories)

sets %>% left_join(inventories) %>%
  group_by(set_num) %>%
  count() %>%
  filter(n > 1)
```



MIDTERMS

```{r}

tset = tibble::tribble(
  ~abb, ~full, ~number, 
  "Jan", "January", 1,
  "Feb", "February", 2,
  "Mar", "March", 3,
  "Apr", "April", 4,
  "May", "May", 5
)

tset %>% mutate(a = match(abb, month.abb))
tset %>% mutate(a = match(full, month.name))


```




